### 2023-12-04 thesis/MSECE_thesis.tex
more thesis edits.

\subsection{Air Traffic Control Complete}\label{sec:atcc}
The Air Traffic Control Complete (ATCC) corpus is a speech corpus consisting of audio recordings with corresponding transcriptions.
ATCC is a collection of three smaller corpora recorded at Dallas-Fort Worth, Logan International, and Washington National airports and
transcribed by current or former air traffic controllers familiar with the respective areas. Audio data was recorded through the placement of Very
High Frequency (VHF) antennae configured to monitor several frequencies at each airport, such as arrival, departure, approach, ground, etc. The types
of frequencies monitored vary by airport, but the presence of departure, approach, and ground frequencies is relatively consistent across subcorpora
\cite{godfrey_air_1994}.
% TODO: data length/size; samples, hours, tokens

\subsection{ATCO2}\label{sec:atco2}
The ATCO2 dataset is a speech corpus that consists of audio communications at Prague and Brno airports, in Czechia, along with corresponding
transcriptions. The speech recordings and transcripts are crowd sourced from volunteers. The dataset was created and labeled, in part, as an English
language detection corpus, however, the labeling also includes transcripts of speech segments which makes it practical for ASR and language model
development in addition to language detection (although in this work it is only used for ASR and LM development)~\cite{szoke_detecting_2021}.
% TODO: data length/size; samples, hours, tokens

\subsection{Air Traffic Control Simulation}\label{sec:atcosim}
The Air Traffic Control Simulation (ATCOSIM) corpus is a speech corpus consisting of audio recordings and corresponding transcriptions.
The data was recorded at the EUROCONTROL Experimental Centre for air traffic control simulation in Bretigny-sur-Orge, France. In this corpus, only the
controllers' voices are included and therefore the transcripts only include the controller's side of each interaction. The audio data was transcribed
by one person, trained according to the guidelines and formatting requirements of the corpus. After all data was transcribed, it was reviewed and
corrected and any remaining problems were reviewed by an operational air traffic controller and resolved \cite{hofbauer_atcosim_2008}.
% TODO: data length/size; samples, hours, tokens

\subsection{ZCU CZ ATC Corpus}\label{sec:zcu_atc}
The ZCU CZ ATC corpus consists of audio recordings and corresponding transcripts in the Czech airspace. Both controller and pilot sides of the
communications are included in this data. Annotations were created by experienced transcribers/annotators and labeled samples were randomly selected
for review during the annotation process. After the dataset was completely labeled, all samples were checked, corrected, and
unified\cite{smidl_air_2019}.
% TODO: data length/size; samples, hours, tokens

@@

A special token is used to represent hesitations when no other context regarding the realization of hesitation is present. If other corpora use a
special token to represent hesitations, it is translated to match the convention in table~\ref{tab:phenomena_conventions}. Partial/cut-off words
are represented with a hyphen (-) at the point where the pronounced word is cut off. For example, if the last syllable of ``approaching'' is cut
mid transmission, it would be represented in the transcript as ``approa-'' or, likewise, if the first syllable is cut off, it would be written as
``-roaching''. Lastly, if there is enough information in the original transcript to indicate the realization of a hesitation, it is represented by
the most appropriate token from ``Filled pauses'' in table~\ref{tab:phenomena_conventions}.

@@

\section{Data Analysis}\label{sec:data_analysis}
% sequence lengths, unique tokens, total number of tokens, total number of samples, breakdown by each dataset
Relevant corpora statistics were observed or calculated and are presented in Table~\ref{tab:corpora_stats}. This includes the number of samples in
the corpus, mean sequence length per sample in the corpus, the number of unique tokens in the corpus, and the total number of tokens in the corpus.
The same statistics were also calculated across corpora, since all four corpora will be combined for training and testing the language and speech
recognition models. Note that this is based on the raw data after loading and unifying the corpora, prior to preprocessing for model training/testing
and outlier removal.

@@

Sci-Kit Learn's \cite{pedregosa_scikit-learn_2011} implementation of the LOF algorithm was used with a value of \(N = 35\) for \(N\) neighbors in the
K-Nearest Neighbors algorithm and Minkowski distance for distance metrics/calculations (in this case, since \(p=2\), Euclidean distance is effectively
used for distance calculations). A total of 47 samples were detected as outliers and removed for analysis and are summarized in table
\ref{tab:outlier_stats}.

@@

Although all of the outliers have higher sequence lengths than the typical sequence lengths of the other samples in the data, going the through
the outlying samples manually and reviewing each reveals that, for the most part, the samples are typical albeit somewhat complex communications.
With the exception of four outliers, all of the samples labeled as outlying by the LOF algorithm were kept in the data.

@@

\begin{equation}\label{eq:ngram_sequence}
    X = \mbox{$w_1$, $w_2$, $w_3$, ..., $w_T$}
\end{equation}

\noindent
the probability of a token, $w_i$, occurring in the sequence can be represented as

\begin{equation}\label{eq:ngram_preceding}
    \begin{gathered}
        P(w_i) = P(w_i|w_{i-1},w_{i-2},w_{i-3},...,w_{i-N}) \\
        P(w_i) = \prod_{j=i-1}^{i-N} P(w_j)
    \end{gathered}
\end{equation}

\noindent
or for the case of succeeding tokens (looking ahead $N$ tokens)

\begin{equation}\label{eq:ngram_succeeding}
    \begin{gathered}
        P(w_i) = P(w_i|w_{i+1},w_{i+2},w_{i+3},...,w_{i+N}) \\
        P(w_i) = \prod_{j=i+1}^{i+N} P(w_j)
    \end{gathered}
\end{equation}

\noindent
where $N$ is the number of preceding (or succeeding) tokens. Since the $N$ preceding tokens\footnote{The most common use case for an N-gram model is
    the preceding tokens.} are needed prior to the estimation of the probability of $w_i$, N-grams are sometimes referred to as left-associative
models (in a left-to-right language).

The probability of a token occurring anywhere in a sequence (referred to as $w_j$ in equations \ref{eq:ngram_preceding} and \ref{eq:ngram_succeeding})
is determined statistically as the number occurrences of a token in a corpus divided by the total number of tokens in the corpus.

@@

\textbf{Next Sentence Prediction}. While MLM conditions the model to represent tokens within sentences, NSP is meant to condition the model to
understand relationships between sentences/sequences to further support downstream tasks. Since BERTs training set consists primarily of literary
works and excerpts from Wikipedia, the  model simply predicts whether two sentences appear together (one after the other) in the original text.
Instead of a single input sequence, the input consists of two sequences separated by a ``separation'' token and beginning with a ``classification''
token, these are represented by \lstinline|[SEP]| and \lstinline|[CLS]|, respectively in Devlin et al.~\cite{devlin_bert_2019}.

For symmetry between pretraining tasks, the special tokens specifying the beginning and end of input sequences are shared between tasks. For both MLM
and NSP objectives, input sequences always start with a classification token and end with a separation token (see figure
\ref{fig:bert_mlm_input_example}). Likewise, NSP inputs always begin with a classification token and each input sequence is concatenated with a
separation token terminating each sequence (see figure \ref{fig:bert_nsp_input_example}).

@@

@@

The tokenization of words with shared roots but different suffixes is shown in table \ref{tab:wordlevel_tokenization_shared_roots_example}. Clearly
all three words are seen as independent and unique words by the tokenizer, since they are mapped to their own integer representations.

@@

The tokenizer used to generate the examples in tables \ref{tab:wordlevel_tokenization_example} and
\ref{tab:wordlevel_tokenization_shared_roots_example} was trained on two sentences, notably lacking the word ``you'll''. Running the tokenizer on the
word ``you'll'' alone results in the output in table \ref{tab:wordlevel_unk_word}. Although the word shares a root with some of the words in the
training data (``you'', ``you're'', and ``you've'') and is semantically very similar, the tokenizer does not recognize the word at all (represented
by the ``[UNK]'' or unknown token). This word would be considered out-of-vocabulary since it did not appear in the training data and thus reveals one
of the major downsides of this algorithm, as mentioned in section \ref{sec:tokenizers}. This problem can be alleviated to some extent by modifying the
preprocessing strategy of the tokenizer to partition contracted words into their component parts e.g.~the word ``you're'' would be partitioned to
\lstinline|[``you'', ``''', ``re'']|, however, it eventually reappears, for example, when using the past, present, and future tenses of a verb
(e.g.~``go'' and ``goes'' in the training data, then encountering the future tense: ``going'').

@@

Table \ref{tab:wordpiece_tokenization_example}, below, shows an example of a WordPiece tokenizer model output on a sentence that was seen during the
training procedure (trained on the same data as the WordLevel tokenizer in Section \ref{sec:wordlevel}). Since this sentence was in the training data,
the segmentation is identical to that of the WordLevel tokenizer in table \ref{tab:wordlevel_tokenization_example}, however, tokenization of similar
words e.g.~words with shared roots or similar contractions are handled differently as in table \ref{tab:wordpiece_tokenization_shared_roots_example}.
Since each word has a shared root word in the contractions (as well as an apostrophe separating the contraction), the tokenizer segments each word and
represents them as groups of small, more common (thus with higher probabilities of occurring) tokens. Additionally, the WordPiece algorithm
effectively handles OOV words. Take the OOV in table \ref{tab:wordpiece_unk_word_similar} with a shared root and similar contraction as those in table
\ref{tab:wordpiece_tokenization_shared_roots_example}, which appear in the training data. Since the base of the contraction, ``you'', occurs in the
training data as well as the apostrophe in the contraction, so the tokenizer segments those parts of the word as tokens. The last part of the word,
``ll'', does not appear at all in the training data, so it is broken down into its component characters which are represented as ``l'' (beginning of
a word) and ``\#\#l'' (segmented part of a word that should be concatenated to the previous token when they are recombined). Comparing this to the
results of the WordLevel tokenizer output on the same word in table \ref{tab:wordlevel_unk_word}, the WordPiece algorithm sidesteps the OOV problem
for words similar to those in the trining corpus and, according to the output of the WordPiece tokenizer in table \ref{tab:wordpiece_unk_word}, for
words unlike those in the training corpus.

@@

Now, testing the BPE algorithm on words that did not appear in the training data, it is clear that the BPE algorithm effectively deals with and
processes OOV words (tables \ref{tab:bpe_unk_similar} and \ref{tab:bpe_unk}), even when no subunits of the word appear in the training data (table
\ref{tab:bpe_unk}).

@@

\textbf{Byte-Level Byte-Pair Encoding}. This is a subset of the BPE algorithm that is also commonly used for NLP tasks such as neural machine
translation, language modeling, and generative tasks (GPT-2 uses a byte-level BPE tokenizer~\cite{radford_language_2019}). As Huggingface's
\lstinline|tokenizers| library implements it, this is a preprocessing step for the tokenizer model that maps all bytes in a string/sentence to their
own unique and visible character. Functionally, byte-level BPE is nearly identical to standard BPE models except that the beginning of word symbol is
actually rendered and represented by a visible character in the tokenized sequence (see table \ref{tab:byte_level_bpe_example}).

@@

The models are trained with the \textit{AdamW} optimizer \cite{loshchilov_decoupled_2019} using the hyperparameters in table \ref{tab:optim_params}
based on the most effective pretraining configuration presented in \cite{devlin_bert_2019} and \cite{liu_roberta_2019}.

@@

\subsection{Performance Metrics}\label{sec:performance_metrics}
As mentioned previously, perplexity and cross-entropy loss (often shortened to loss) are used to track the performance of the model on the training
and validation sets, per epoch, while training. Cross-entropy loss is defined by the following equation:

@@

\noindent
where $x$ is a $512 \times N$ vector representing model inputs (a sequence of tokens) and $y$ is a $1 \times N$ vector representing the target/label
for each sample in the batch. $N$ is the number of samples in the batch, $l_n$ is the loss for an individual sample, $n$, in the batch, and $C$ is the
number of class labels. For the purpose of language modeling, the set of class labels is equivalent to the vocabulary of the tokenizer (and thus the
language model). Numerical labels will correspond to token IDs. The overall loss is then found by taking the mean of the loss over the batch, likewise
for an epoch, the loss is found by taking the average of each batch in the epoch.

Perplexity is related to the probability of a token appearing in a sequence. It is used as a measure of how well a language model can predict tokens
in a sequence and is defined by the equation below:

@@

The perplexity of the models on the training data yields interesting results (figure \ref{fig:training_ppl}). Although the pretrained models have a
more stable curve over time, the models trained from scratch result in a significantly lower perplexity. Note the difference in the scales on the
y-axis in figure \ref{fig:training_ppl}; the final perplexity value of each language model trained from scratch is at least two orders of magnitude
lower than that of the pretrained models. At the end of training, the BERT and RoBERTa models trained from scratch had perplexity values of
$0.52\times 10^7$ and $0.16\times 10^7$, respectively, whereas the pretrained models reached perplexity values of $0.29\times 10^9$ and
$0.17\times 10^9$, for BERT and RoBERTa, respectively. This indicates that best language model performance may be achieved by pretraining models on
in-domain data, then continuing to fine-tune on specific downstream tasks. This is a direct contradiction of the current recommended practice of using
one of the pretrained checkpoints from Devlin et al.~\cite{devlin_bert_2019} and Liu et al.~\cite{liu_roberta_2019} then fine-tuning on specific
downstream tasks. This is further upheld by the results of the model performance on the test partition of the corpus in table
\ref{tab:pretraining_results}, below. As with the training data, the perplexity of the language models trained from scratch over the test partition of
the corpus is significantly lower than that of the pretrained language models. The loss of each model over the test set is somewhat high, but they are
around the same values for each model, which is consistent with the loss statistics during training. A higher loss value on unseen (test) data is
normal and expected. The BERT model trained from scratch had the best performance among the models tested despite both RoBERTa models training better
within their model classes. Within the class of pretrained models, however, the RoBERTa model still technically performed better than the BERT model
(values shown in table \ref{tab:pretraining_results} were rounded to three significant figures) with a perplexity value of $0.007\times 10^9$ compared
to BERT's $0.014\times 10^9$.

@@

\textbf{Mixed regional data sources}. As described in Section \ref{sec:data_source}, the data aggregated for use here is comes from several different
regions, namely international airports in the US \cite{godfrey_air_1994} and international and domestic airports in Europe
\cite{smidl_air_2019,hofbauer_atcosim_2008,szoke_detecting_2021}. During data processing it was ensured that only English transcriptions were included
in the corpus, however, proper nouns from each region will be present in the data regardless. Additionally, dialectical differences and varying levels
of English proficiency will affect the structure of token sequences across regions. All of these facets combined may be contributing to the high
perplexity of the language models relative to other works. It should be noted that the varied regional sources of data are not necessarily a bad
thing. Depending on the application and desired robustness of the model, the ability to interpret international sources and contexts can be a
desirable trait for model selection, however, it will require significantly more data and proportionally balanced regional contexts within the corpus
in order for the pretraining process to be effective for downstream tasks.

Up to this point we have followed a procedure for studying the effectiveness of the models that is somewhat symmetrical to that of BERT and RoBERTa,
so at this point the models should be trained for downstream tasks in order to further measure the effectiveness of the pretraining procedures.
In the aviation domain and for this particular data, that would mean training for tasks such as callsign identification, speaker identification,
flight phase classification, clearance recognition, etc., however, labels for these tasks are not present in the corpora used here. Corpora with
relevant labels for downstream tasks could not be obtained within an appropriate amount of time to run and include the experiments in this work.
Some potential algorithms were developed in an attempt to mine labels from the corpora aggregated for this work which will be expanded upon in section
\ref{sec:future_work}. For these reasons, testing model performance and effectiveness on downstream tasks must be left for future work.

\section{Conclusion}\label{sec:conclusion}
Four aviation English corpora are aggregated and unified for use in neural language modeling. A masked language modeling objective is used pretrain
two classes of BERT and RoBERTa models (models trained from scratch and pretrained models transfer learned to aviation English) on the aggregated
data.

@@

The results from the experiments in this work indicate that, for natural language processing tasks in the aviation domain, models pretrained from
scratch are more effective than transfer learning pretrained models. Thus, models trained from scratch should fine-tuned and used for downstream tasks
over pretrained models. The most significant limiting factor to the performance of the models is the small volume of aviation English data available
for pretraining and downstream task use.

@@

\newpage
\section{Appendix A: Examples of Samples Detected as Outliers}

>>>>

\subsection{Air Traffic Control Complete}\label{sec:atcc}
The Air Traffic Control Complete (ATCC) corpus is a speech corpus consisting of audio recordings with corresponding transcriptions.
ATCC is a collection of three smaller corpora recorded at Dallas-Fort Worth, Logan International, and Washington National airports and
transcribed by current or former air traffic controllers familiar with the respective areas. Audio data was recorded through the placement of Very
High Frequency (VHF) antennae configured to monitor several frequencies at each airport, such as arrival, departure, approach, ground, etc. The types
of frequencies monitored vary by airport, but the presence of departure, approach, and ground frequencies is relatively consistent across subcorpora
\cite{godfrey_air_1994}. See Table \ref{tab:atcc_stats} for corpus-specific statistics.

\begin{table}
    \centering
    \begin{tabular}{l r}
        \toprule
        \multicolumn{2}{c}{ATCC Corpus Properties} \\
        \midrule
        Samples              & 9,556               \\
        Audio (hours)        & 72.48               \\
        Mean Sequence Length & 11.90               \\
        Std.~Sequence Length & 7.22                \\

        Total Tokens         & 355,485             \\
        Unique Tokens        & 2,209               \\
        \bottomrule
    \end{tabular}
    \caption{Data statistics for the Air Traffic Control Complete (ATCC) corpus.}
    \label{tab:atcc_stats}
\end{table}

\subsection{ATCO2}\label{sec:atco2}
The ATCO2 dataset is a speech corpus that consists of audio communications at Prague and Brno airports, in Czechia, along with corresponding
transcriptions. The speech recordings and transcripts are crowd sourced from volunteers. The dataset was created and labeled, in part, as an English
language detection corpus, however, the labeling also includes transcripts of speech segments which makes it practical for ASR and language model
development in addition to language detection (although in this work it is only used for ASR and LM development)~\cite{szoke_detecting_2021}.
See Table \ref{tab:atco2_stats} for corpus-specific statistics.

\begin{table}
    \centering
    \begin{tabular}{l r}
        \toprule
        \multicolumn{2}{c}{ATCO2 Corpus Properties} \\
        \midrule
        Samples              & 874                  \\
        Audio (hours)        & 1.10                 \\
        Mean Sequence Length & 12.28                \\
        Std.~Sequence Length & 5.78                 \\
        Total Tokens         & 10,733               \\
        Unique Tokens        & 786                  \\
        \bottomrule
    \end{tabular}
    \caption{Data statistics for the ATCO2 corpus.}
    \label{tab:atco2_stats}
\end{table}

\subsection{Air Traffic Control Simulation}\label{sec:atcosim}
The Air Traffic Control Simulation (ATCOSIM) corpus is a speech corpus consisting of audio recordings and corresponding transcriptions.
The data was recorded at the EUROCONTROL Experimental Centre for air traffic control simulation in Bretigny-sur-Orge, France. In this corpus, only the
controllers' voices are included and therefore the transcripts only include the controller's side of each interaction. The audio data was transcribed
by one person, trained according to the guidelines and formatting requirements of the corpus. After all data was transcribed, it was reviewed and
corrected and any remaining problems were reviewed by an operational air traffic controller and resolved \cite{hofbauer_atcosim_2008}. See Table
\ref{tab:atcosim_stats} for corpus-specific statistics.

\begin{table}
    \centering
    \begin{tabular}{l r}
        \toprule
        \multicolumn{2}{c}{ATCOSIM Corpus Properties} \\
        \midrule
        Samples              & 9,556                  \\
        Audio (hours)        & 10.69                  \\
        Mean Sequence Length & 11.29                  \\
        Std.~Sequence Length & 4.17                   \\
        Total Tokens         & 107,881                \\
        Unique Tokens        & 827                    \\
        \bottomrule
    \end{tabular}
    \caption{Data statistics for the ATCOSIM corpus.}
    \label{tab:atcosim_stats}
\end{table}

\subsection{ZCU CZ ATC Corpus}\label{sec:zcu_atc}
The ZCU CZ ATC corpus consists of audio recordings and corresponding transcripts in the Czech airspace. Both controller and pilot sides of the
communications are included in this data. Annotations were created by experienced transcribers/annotators and labeled samples were randomly selected
for review during the annotation process. After the dataset was completely labeled, all samples were checked, corrected, and
unified\cite{smidl_air_2019}. See Table \ref{tab:zcu_cz_atc_stats} for corpus-specific statistics.

\begin{table}
    \centering
    \begin{tabular}{l r}
        \toprule
        \multicolumn{2}{c}{ZCU CZ ATC Corpus Properties} \\
        \midrule
        Samples              & 14,435                    \\
        Audio (hours)        & 20.58                     \\
        Mean Sequence Length & 10.05                     \\
        Std.~Sequence Length & 5.50                      \\
        Total Tokens         & 145,107                   \\
        Unique Tokens        & 3,100                     \\
        \bottomrule
    \end{tabular}
    \caption{Data statistics for ZCU CZ ATC corpus.}
    \label{tab:zcu_cz_atc_stats}
\end{table}

@@

A special token is used to represent hesitations when no other context regarding the realization of hesitation is present. If other corpora use a
special token to represent hesitations, it is translated to match the convention in Table \ref{tab:phenomena_conventions}. Partial/cut-off words
are represented with a hyphen (-) at the point where the pronounced word is cut off. For example, if the last syllable of ``approaching'' is cut
mid transmission, it would be represented in the transcript as ``approa-'' or, likewise, if the first syllable is cut off, it would be written as
``-roaching''. Lastly, if there is enough information in the original transcript to indicate the realization of a hesitation, it is represented by
the most appropriate token from ``Filled pauses'' in Table \ref{tab:phenomena_conventions}.

@@

\section{Data Analysis}\label{sec:data_analysis}
% sequence lengths, unique tokens, total number of tokens, total number of samples, breakdown by each dataset
Relevant corpora statistics were observed or calculated and are presented in Table \ref{tab:corpora_stats}. This includes the number of samples in
the corpus, mean sequence length per sample in the corpus, the number of unique tokens in the corpus, and the total number of tokens in the corpus.
The same statistics were also calculated across corpora, since all four corpora will be combined for training and testing the language and speech
recognition models. Note that this is based on the raw data after loading and unifying the corpora, prior to preprocessing for model training/testing
and outlier removal.

@@

Sci-Kit Learn's \cite{pedregosa_scikit-learn_2011} implementation of the LOF algorithm was used with a value of \(N = 35\) for \(N\) neighbors in the
K-Nearest Neighbors algorithm and Minkowski distance for distance metrics/calculations (in this case, since \(p=2\), Euclidean distance is effectively
used for distance calculations). A total of 47 samples were detected as outliers and removed for analysis and are summarized in Table
\ref{tab:outlier_stats}.

@@

Although all of the outliers have higher sequence lengths than the typical sequence lengths of the other samples in the data, going the through
the outlying samples manually and reviewing each reveals that, for the most part, the samples are typical albeit somewhat complex communications.
With the exception of four outliers, all of the samples labeled as outlying by the LOF algorithm were kept in the data. Examples of the outliers are
listed in Appendix A in Section \ref{sec:appendix_a}, including the sample removed.

@@

\begin{equation}\label{eq:ngram_sequence}
    X = \mbox{$w_1$, $w_2$, $w_3$, ..., $w_T$}\\
\end{equation}

\noindent
the probability of tokens occurring in order, in a sequence can be represented using the chain-rule of probability:

\begin{equation}
    \begin{gathered}
        P(X) = P(w_1) P(w_2|w_1) P(w_3|w_2, w_1) ... P(w_T|w_{T-1}, w_{T-2}, ..., w_1)\\
        P(X) = P(w_i)\prod_{i=2}^{T} P(w_i|w_{i-1}, w_{i-2}, ..., w_1)
    \end{gathered}
\end{equation}

\noindent
Except for the first token in the sequence, the probability of a token occurring at a specific point in the sequence depends on the tokens preceding it
\footnote{The most common use case for an N-gram model is the preceding tokens.}. Due to this fact, N-grams are sometimes referred to as
left-associative models (in a left-to-right language).

The probability of a token occurring at all is determined statistically as the number occurrences of a token in a corpus divided by the total number
of tokens in the corpus.

@@

\textbf{Next Sentence Prediction}. While MLM conditions the model to represent tokens within sentences, NSP is meant to condition the model to
understand relationships between sentences/sequences to further support downstream tasks. Since BERTs training set consists primarily of literary
works and excerpts from Wikipedia, the  model simply predicts whether two sentences appear together (one after the other) in the original text,
making this a binary classification task as opposed to MLM which is a multi-class classification task. Instead of a single input sequence, the input
consists of two sequences separated by a ``separation'' token and beginning with a ``classification'' token, these are represented by
\lstinline|[SEP]| and \lstinline|[CLS]|, respectively in Devlin et al.~\cite{devlin_bert_2019}.

For symmetry between pretraining tasks, the special tokens specifying the beginning and end of input sequences are shared between tasks. For both MLM
and NSP objectives, input sequences always start with a classification token and end with a separation token (see Figure
\ref{fig:bert_mlm_input_example}). Likewise, NSP inputs always begin with a classification token and each input sequence is concatenated with a
separation token terminating each sequence (see Figure \ref{fig:bert_nsp_input_example} and equations (\ref{eq:bert_mlm_input_example}) and
(\ref{eq:bert_nsp_input_example})).

\begin{equation}\label{eq:bert_mlm_input_example}
    \mbox{[CLS], the, quick, brown, fox, [SEP]}
\end{equation}

@@

\begin{equation}\label{eq:bert_nsp_input_example}
    \mbox{[CLS], the, quick, brown, fox, [SEP], jumps, over, the, [SEP]}
\end{equation}

@@

The tokenization of words with shared roots but different suffixes is shown in Table \ref{tab:wordlevel_tokenization_shared_roots_example}. Clearly
all three words are seen as independent and unique words by the tokenizer, since they are mapped to their own integer representations.

@@

The tokenizer used to generate the examples in tables \ref{tab:wordlevel_tokenization_example} and
\ref{tab:wordlevel_tokenization_shared_roots_example} was trained on two sentences, notably lacking the word ``you'll''. Running the tokenizer on the
word ``you'll'' alone results in the output in Table \ref{tab:wordlevel_unk_word}. Although the word shares a root with some of the words in the
training data (``you'', ``you're'', and ``you've'') and is semantically very similar, the tokenizer does not recognize the word at all (represented
by the ``[UNK]'' or unknown token). This word would be considered out-of-vocabulary since it did not appear in the training data and thus reveals one
of the major downsides of this algorithm, as mentioned in section \ref{sec:tokenizers}. This problem can be alleviated to some extent by modifying the
preprocessing strategy of the tokenizer to partition contracted words into their component parts e.g.~the word ``you're'' would be partitioned to
\lstinline|[``you'', ``''', ``re'']|, however, it eventually reappears, for example, when using the past, present, and future tenses of a verb
(e.g.~``go'' and ``goes'' in the training data, then encountering the future tense: ``going'').

@@

Table \ref{tab:wordpiece_tokenization_example}, below, shows an example of a WordPiece tokenizer model output on a sentence that was seen during the
training procedure (trained on the same data as the WordLevel tokenizer in Section \ref{sec:wordlevel}). Since this sentence was in the training data,
the segmentation is identical to that of the WordLevel tokenizer in Table \ref{tab:wordlevel_tokenization_example}, however, tokenization of similar
words e.g.~words with shared roots or similar contractions are handled differently as in Table \ref{tab:wordpiece_tokenization_shared_roots_example}.
Since each word has a shared root word in the contractions (as well as an apostrophe separating the contraction), the tokenizer segments each word and
represents them as groups of small, more common (thus with higher probabilities of occurring) tokens. Additionally, the WordPiece algorithm
effectively handles OOV words. Take the OOV in Table \ref{tab:wordpiece_unk_word_similar} with a shared root and similar contraction as those in Table
\ref{tab:wordpiece_tokenization_shared_roots_example}, which appear in the training data. Since the base of the contraction, ``you'', occurs in the
training data as well as the apostrophe in the contraction, so the tokenizer segments those parts of the word as tokens. The last part of the word,
``ll'', does not appear at all in the training data, so it is broken down into its component characters which are represented as ``l'' (beginning of
a word) and ``\#\#l'' (segmented part of a word that should be concatenated to the previous token when they are recombined). Comparing this to the
results of the WordLevel tokenizer output on the same word in Table \ref{tab:wordlevel_unk_word}, the WordPiece algorithm sidesteps the OOV problem
for words similar to those in the trining corpus and, according to the output of the WordPiece tokenizer in Table \ref{tab:wordpiece_unk_word}, for
words unlike those in the training corpus.

@@

Now, testing the BPE algorithm on words that did not appear in the training data, it is clear that the BPE algorithm effectively deals with and
processes OOV words (tables \ref{tab:bpe_unk_similar} and \ref{tab:bpe_unk}), even when no subunits of the word appear in the training data (Table
\ref{tab:bpe_unk}).

@@

\textbf{Byte-Level Byte-Pair Encoding}. This is a subset of the BPE algorithm that is also commonly used for NLP tasks such as neural machine
translation, language modeling, and generative tasks (GPT-2 uses a byte-level BPE tokenizer~\cite{radford_language_2019}). As Huggingface's
\lstinline|tokenizers| library implements it, this is a preprocessing step for the tokenizer model that maps all bytes in a string/sentence to their
own unique and visible character. Functionally, byte-level BPE is nearly identical to standard BPE models except that the beginning of word symbol is
actually rendered and represented by a visible character in the tokenized sequence (see Table \ref{tab:byte_level_bpe_example}).

@@

The models are trained with the \textit{AdamW} optimizer \cite{loshchilov_decoupled_2019} using the hyperparameters in Table \ref{tab:optim_params}
based on the most effective pretraining configuration presented in \cite{devlin_bert_2019} and \cite{liu_roberta_2019}.

@@

\subsection{Performance Metrics}\label{sec:performance_metrics}
\textbf{Cross-Entropy Loss}. As mentioned previously, perplexity and cross-entropy loss (often shortened to loss) are used to track the performance of
the model on the training and validation sets, per epoch, while training. Cross-entropy loss is defined by the following equation:

@@

\noindent
where $x$ is a $512 \times N$ vector representing model inputs (a sequence of tokens) and $y$ is a $1 \times N$ vector representing the target/label
for each sample in the batch. $L$ is an $N \times 1$ vector containing the losses for each sample in the batch, $N$ is the number of samples in the
batch, $l_n$ is the loss for an individual sample, $n$, in the batch, and $C$ is the number of class labels. For the purpose of language modeling, the
set of class labels is equivalent to the vocabulary of the tokenizer (and thus the language model). Numerical labels will correspond to token IDs. The
overall loss is then found by taking the mean of the loss over the batch, likewise for an epoch, the loss is found by taking the average of each batch
in the epoch.

\textbf{Perplexity}. Perplexity is related to the probability of a token appearing in a sequence. It is used as a measure of how well a language model
can predict tokens in a sequence and is defined by the equation below:

@@

The perplexity of the models on the training data yields interesting results (Figure \ref{fig:training_ppl}). Although the pretrained models have a
more stable curve over time, the models trained from scratch result in a significantly lower perplexity. Note the difference in the scales on the
y-axis in Figure \ref{fig:training_ppl}; the final perplexity value of each language model trained from scratch is at least two orders of magnitude
lower than that of the pretrained models. At the end of training, the BERT and RoBERTa models trained from scratch had perplexity values of
$0.52\times 10^7$ and $0.16\times 10^7$, respectively, whereas the pretrained models reached perplexity values of $0.29\times 10^9$ and
$0.17\times 10^9$, for BERT and RoBERTa, respectively. This indicates that best language model performance may be achieved by pretraining models on
in-domain data, then continuing to fine-tune on specific downstream tasks. This is a direct contradiction of the current recommended practice of using
one of the pretrained checkpoints from Devlin et al.~\cite{devlin_bert_2019} and Liu et al.~\cite{liu_roberta_2019} then fine-tuning on specific
downstream tasks. This is further upheld by the results of the model performance on the test partition of the corpus in Table
\ref{tab:pretraining_results}, below. As with the training data, the perplexity of the language models trained from scratch over the test partition of
the corpus is significantly lower than that of the pretrained language models. The loss of each model over the test set is somewhat high, but they are
around the same values for each model, which is consistent with the loss statistics during training. A higher loss value on unseen (test) data is
normal and expected. The BERT model trained from scratch had the best performance among the models tested despite both RoBERTa models training better
within their model classes. Within the class of pretrained models, however, the RoBERTa model still technically performed better than the BERT model
(values shown in Table \ref{tab:pretraining_results} were rounded to three significant figures) with a perplexity value of $0.007\times 10^9$ compared
to BERT's $0.014\times 10^9$.

@@

\textbf{Mixed regional data sources}. As described in Section \ref{sec:data_source}, the data aggregated for use here comes from several different
regions, namely international airports in the US \cite{godfrey_air_1994} and international and domestic airports in Europe
\cite{smidl_air_2019,hofbauer_atcosim_2008,szoke_detecting_2021}. During data processing it was ensured that only English transcriptions were included
in the corpus, however, proper nouns from each region will be present in the data regardless. Additionally, dialectical differences and varying levels
of English proficiency will affect the structure of token sequences across regions. All of these facets combined may be contributing to the high
perplexity of the language models relative to other works. It should be noted that the varied regional sources of data are not necessarily a bad
thing. Depending on the application and desired robustness of the model, the ability to interpret international sources and contexts can be a
desirable trait for model selection, however, it will require significantly more data and proportionally balanced regional contexts within the corpus
in order for the pretraining process to be effective for downstream tasks.

Up to this point we have followed a procedure for studying the effectiveness of the models that is somewhat symmetrical to that of BERT and RoBERTa,
so at this point the models should be trained for downstream tasks in order to further measure the effectiveness of the pretraining procedures.
In the aviation domain and for this particular data, that would mean training for tasks such as callsign identification, speaker identification,
flight phase classification, clearance recognition, etc. However, labels for these tasks are not present in the corpora used here. Corpora with
relevant labels for downstream tasks could not be obtained within an appropriate amount of time to run and include the experiments in this work.
Some potential algorithms were developed in an attempt to mine labels from the corpora aggregated for this work which will be expanded upon in section
\ref{sec:future_work}. For these reasons, testing model performance and effectiveness on downstream tasks must be left for future work.

\section{Conclusion}\label{sec:conclusion}
Four aviation English corpora are aggregated and unified for use in neural language modeling. Masked language modeling is then used pretrain two
classes of BERT and RoBERTa models (models trained from scratch and pretrained models transfer learned to aviation English).

@@

The results from the experiments in this work indicate that, for natural language processing tasks in the aviation domain, models pretrained from
scratch are more effective than transfer learning pretrained models. Thus, models trained from scratch should be fine-tuned and used for downstream
tasks over pretrained models. The most significant limiting factor to the performance of the models is the small volume of aviation English data
available for pretraining and downstream task use.

@@

\newpage
\section{Appendix A: Examples of Samples Detected as Outliers}
The outlying sample that was removed from the corpus is below\footnote{The preprocessing effects on the string, capitalization, punctuation, etc.~are left
    unmodified for display purposes}:

\begin{quote}
    uh uh have a nice evening thank you bye bye ciao yes to the irish pub you know it uh you have to ask i have no idea where it is i'm just walking
    behind the others uh no uh but you come in front and you ask and he will tell you but i really don't know i don't even know the address okay it's
    in paris definitely okay bye
\end{quote}

This sample has a sequence length of 71 tokens, well above the mean of the aggregated corpora. The decision was made to remove this sample, because
it has little to no relevance to aviation communications, especially the ones being examined by the model in this thesis.

A few other samples, detected as outliers, but not removed are shown in Table \ref{tab:outlier_examples}.
